### 2024-03-26 doc/sections/mav-semantics.tex
[ tex ] various

\subsection{$*$-autonomous and Duoidal Partial Orders}

@@

@@

\begin{remark}
  The structure of a $*$-autonomous partial order has a number of
  immediate consequences, but we leave description of these until
  after the definition of MAV-algebra below. \bob{because there are
    even more consequences in the presence of additive and duoidal
    structure}
\end{remark}

@@

\bob{need to define \emph{entropy} and \emph{entropic} somewhere}

@@

\begin{definition}
  An \emph{MAV-algebra} is a structure
  $(A, \leq, \otimes, ;, I, \lnot)$ with the following properties:
  \begin{enumerate}
  \item $(A, \leq, \otimes, I, \lnot)$ is $*$-autonomous and satisfies \emph{mix}.
  \item $(A, \leq, ;, I)$ is a pomonoid.
  \item $;$ is self dual: $\lnot (x ; y) = (\lnot x); (\lnot y)$.
  \item $(\otimes, I)$ is duoidal over $(;, I)$.
  \item $(A, \leq)$ has binary meets, which we write as $x \with y$.
  \end{enumerate}
\end{definition}

@@

The Lower Set adds residuals to the pomonoids of an MAV-frame, so the
Chu construction below will allow us to add duals to interpret all of
the multiplicative subsystem of MAV. To incorporate the additives in
such a way that we will be able to derive the proof normalisation
property, we need to refine lower sets to ones that are closed under
$+$-combinations.

@@

\begin{proposition}\label{prop:closed-monoid-distrib}
  For a commutative pomonoid $(\bullet, i)$ on $(A, \leq)$ that
  distributes over $+$
  ($(x \bullet y) + z = (x + z) \bullet (y + z)$), then
  $F \ClosedDay{\bullet} G = \alpha(U F \Day{\bullet} U G)$ and
  $\ClosedDay{j} = \alpha(\Day{j})$ define a residuated commutative
  pomonoid on $\ClosedLowerSet{A}$.
\end{proposition}

>>>>

\bob{Short introduction to the section here}

\subsection{$*$-autonomous and Duoidal Structure on Partial Orders}

@@

The definition of $*$-autonomous category is due to Barr
\cite{Barr_1979}. For our purposes, we need the partial order
analogue, also called a CL algebra by Troelstra
\cite{Troelstra92:lll}.

@@

\begin{remark}
  The structure of a $*$-autonomous partial order has a number of
  immediate consequences, but we leave description of these until
  after the definition of MAV-algebra in Definition
  \ref{defn:mav-algebra}.
\end{remark}

@@

\bob{need to define \emph{entropy} and \emph{entropic} somewhere. I
  think we can have that one pomonoid is \emph{entropic} over another,
  and the structure of having two monoids where one is entropic over
  the other is called duoidal structure.}

@@

\begin{definition}\label{defn:mav-algebra}
  An \emph{MAV-algebra} is a structure
  $(A, \leq, \otimes, ;, I, \lnot)$ with the following properties:
  \begin{enumerate}
  \item $(A, \leq, \otimes, I, \lnot)$ is $*$-autonomous and satisfies \emph{mix}.
  \item $(A, \leq, ;, I)$ is a pomonoid.
  \item $;$ is self dual: $\lnot (x ; y) = (\lnot x); (\lnot y)$.
  \item $(\otimes, I)$ is duoidal over $(;, I)$.
  \item $(A, \leq)$ has binary meets, which we write as $x \with y$.
  \end{enumerate}
\end{definition}

@@

In the phase semantics, formulas are interpreted as elements that are
fixed points of a closure operator defined by a double negation with
respect to the monoid on the original frame. This closure operator
generates a partial order of ``facts'' whose meets and joins exactly
correspond to the syntactic ones when the original monoid is derived
from the proofs. As we mentioned in the introduction, the presence of
a self-dual operator in BV means that we cannot use double negation
closure, and we have to proceed more deliberately to preserve
join-like structure in an MAV-frame when building MAV-algebras.

We do this by defining closed lower sets as those that are closed
under finite $+$-combinations of their elements.

Lower sets add residuals to the pomonoids of an MAV-frame, so the Chu
construction below will allow us to add duals to interpret all of the
multiplicative subsystem of MAV. To incorporate the additives in such
a way that we will be able to derive the proof normalisation property,
we need to refine lower sets to ones that are closed under
$+$-combinations.

@@

\begin{proposition}\label{prop:closed-monoid-distrib}
  For a commutative pomonoid $(\bullet, i)$ on $(A, \leq)$ that
  distributes over $+$
  ($(x + y) \bullet z \leq (x \bullet z) + (y \bullet z)$), then
  $F \ClosedDay{\bullet} G = \alpha(U F \Day{\bullet} U G)$ and
  $\ClosedDay{j} = \alpha(\Day{j})$ define a residuated commutative
  pomonoid on $\ClosedLowerSet{A}$.
\end{proposition}
